IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/concrete.tex
Structure up the text.
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

To obtain an isomorphism with our single heap semantics, we need to mimic the
behavior of several heaps with a single actual heap.
The interesting cases are when we sandbox
an expression, and when messages are sent and received.
The rule for sandboxing is
parameterized by the strategy |klone| (cf.\ Section~\ref{sec:retrofit}),
which defines what heap the new task
should execute with.  We have motivated two choices, namely the identity (to
effectively clone the heap), or to start with an empty heap.

{\color{red}
In the first choice, the concrete semantics should essentially implement several
heaps into one. For that, it creates a copy of everything reachable from the
forked expression (not just the values for the memory addresses in the
expression \Red{Ale:why not only the addresses?}), and replace all addresses
correspondingly.  behavior of sending a memory address now depends on whether
that address existed at the time the receiving task was forked; if it did, then
the address should be rewritten to the new one, otherwise to a fresh one.

In the latter case, this effectively means that the concrete semantics must
change the forked expression |ie| in such a way that all addresses get rewritten
to fresh addresses that are guaranteed to not occur in the global heap. As a
consequence, existing addresses in the forked expression |ie| are no longer
valid and the task gets stuck (and then removed by \textsc{I-noStep}).
%To get the isomorphic behavior in the concrete semantics with a
%single heap, we would have to change the forked expression |ie| in
%such a way that all addresses get rewritten to fresh addresses
%that are guaranteed to not occur in the global heap.  
Similarly, sending a memory address should take a similar approach, as addresses
for sent references will not exist in the receiving task.
}

{\color{blue}
In the latter case, this effectively means that existing addresses
in the forked expression |ie| are no longer valid and the
task gets stuck (and then removed by \textsc{I-noStep}).
To get the isomorphic behavior in the concrete semantics with a
single heap, we would have to change the forked expression |ie| in
such a way that all addresses get rewritten to fresh addresses
that are guaranteed to not occur in the heap.  Similarly,
sending a memory address should take a similar approach, as addresses
for sent references will not exist in the receiving task.

If |klone| is the identity function, then the abstract semantics
effectively clone the heap on a fork.  To get an isomorphic
behavior in the concrete, we would have to create a copy of everything
reachable from the forked expression (not just the values for
the memory addresses in the expression), and replace all addresses
correspondingly.  Even worse, the behavior of sending a memory address
now depends on whether that address existed at the time the receiving
task was forked;  if it did, then the address should be rewritten to the
new one, otherwise to a fresh address.

Such a complicated behavior is contradictory to our initial motivation
of implementing a single heap for efficiency.  But even the
behavior of the simpler case where |klone| returns an empty heap is
undesirable:  We rewrite addresses to fresh addresses, which will
result in a stuck task whenever such an address is dereferenced.
}

@@

As a better solution we could forbid forked expressions as well
as messages sent to other tasks to contain memory addresses in the
first place.  In a statically typed language, the type system could
prevent this from happening, and we talk more about how our
approach integrates with typed languages in
Section~\ref{sec:extensions:types}.
In a dynamically type languages such as |targetLangML|, we might
restrict the transition for |fork| and |send| to only allow expressions
without memory addresses.

Again, we are modifying the IFC language semantics in a non-trivial way,
which begs the question whether non-interference is preserved.
Restricting the transitions in an information flow control
system in general does not preserve non-interference.  Intuitively
if the restriction depends on secret data, then a public thread
can observe if some other task terminates or not, and by that obtain
information about the secret data that was used to restrict the
transition.
With this in mind, we require semantic rules to get restricted only
based on information observable by the task triggering them.
This ensures that non-interference is preserved, as the
restriction does not depend on confidential information.
Below, we give the formal definition of this condition for the
abstract IFC language |specLang alpha targetLang|.

>>>>

Unfortunately, an isomorphism is often too strong of a requirement.
To obtain an isomorphism with our single heap semantics, we need to mimic the
behavior of several heaps with a single actual heap.
The interesting cases are when we sandbox
an expression and when messages are sent and received.
The rule for sandboxing is
parameterized by the strategy |klone| (cf.\ Section~\ref{sec:retrofit}),
which defines what heap the new task
should execute with.  We have considered two choices:

\begin{itemize}
    \item When we sandbox into an empty heap, existing addresses
in the forked expression are no longer valid and the
task will get stuck (and then removed by \textsc{I-noStep}).
Thus, we must rewrite the forked expression so that
all addresses point to fresh addresses
guaranteed to not occur in the heap.  Similarly,
sending a memory address should be rewritten.

\item When we clone the heap, we have to create a copy of everything
reachable from the forked expression (not just the values for
the memory addresses in the expression), and replace all addresses
correspondingly.  Even worse, the behavior of sending a memory address
now depends on whether that address existed at the time the receiving
task was forked;  if it did, then the address should be rewritten to the
new one, otherwise to a fresh address.
\end{itemize}

This complicated behavior is counter to our initial motivation
of implementing a single heap for efficiency.

@@

A better solution is to forbid forked expressions as well
as messages sent to other tasks to contain memory addresses in the
first place.  In a statically typed language, the type system could
prevent this from happening, and we talk more about how our
approach integrates with typed languages in
Section~\ref{sec:extensions:types}.
In a dynamically type languages such as |targetLangML|, we might
restrict the transition for |fork| and |send| to only allow expressions
without memory addresses.

While this sounds plausible, it is worth noting that we are modifying the IFC language semantics,
which raises the question whether non-interference is preserved.
This question can be subtle: it is easy to remove a transition from
a language and invalidate TSNI.  Intuitively
if the restriction depends on secret data, then a public thread
can observe if some other task terminates or not, and by that obtain
information about the secret data that was used to restrict the
transition.
With this in mind, we require semantic rules to get restricted only
based on information observable by the task triggering them.
This ensures that non-interference is preserved, as the
restriction does not depend on confidential information.
Below, we give the formal definition of this condition for the
abstract IFC language |specLang alpha targetLang|.
